---
title: "portOptimizer Example: addTermToObjective"
author: "Rex Macey"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro to portOptimizer using addTermToObjective to build the objective function.}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(magrittr)
library(knitr)
library(kableExtra)
library(tibble)
library(portOptimizer)

```

# Overview    
The `portOptimizer` package's engine is the `maximize_objective` function.  True to its name, it maximizes the value of a function.  Consider the following generalization of an objective function which has **terms**.  Each **term** in the objective function is a function of x which is a vector of portfolio weights. For example, we might have    
* f1.value = f1(x, params1) # this could be expected return    
* f2.value = f2(x, params2) # this could be expected risk    
* ... more functions    
* fn.value = fn(x, paramsN) # this could be the number of non-zero weights.    

The objective function might look something like the following.  In this case the first term might be called an `objective` and the others `contraints`. The constraints are aggregated to calculate the penalty.    
= objective - penalty     
= f1.value -     
  (ifelse(f2.value >= f2.min, 0, (f2 - min1) ^ 2) +    
  ifelse(f2.value <= f2.max, 0, (f2 - max1) ^ 2) +    
  ... more terms here +    
  ifelse(fn.value >= fn.min, 0, (fn - minN) ^ 2) +    
  ifelse(fn.value <= fn.max, 0, (fn - maxN) ^ 2)) * 25000000000    
  
You may use any objective function so long as the first parameter is x and the function returns a scalar value.  The `portOptimizer` package includes the `add_term_to_objective` function to help create an objective function.

 `add_term_to_objective` has the following parameters     
* `name`: A description.  This might not be needed, but might help if we want to provide a breakdown of each term's contribution to the objective function. 
* `type` Either 'maximize', 'minimize', or 'contraint'. If a constraint, then the `min` and/or `max` should come into play.
* `funName`: The function (or maybe its name). This would be the f1, f2, ..., fn which might be to calculate portfolio return.
* `params`: The (not x) parameters needed by `funName`.  All functions will need the weights (x). We assume that's the first parameter and it should be ommitted from `params`.   
* `min`: Minimum allowed value for constraints. Default is -Inf.  Ignored if -Inf.
* `max`: Maximum allowed value for constraints. Default is Inf. Ignored if Inf.

# Description of the Objedtive Function    
In the following objective definition: 
* MaxReturn: maximize the expected returns subject to the following constraints
* Risk: risk should not be more than the risk of benchX. Note how the max is set to the risk of the benchmark
* TrackingError: tracking error should not be more the 0.03
* NAssets: number of assets should be between 3 and 10. Note the empty params list, and use of both min and max.
* MinNonZeroWt: minimum non-zero weight should be at least 0.03 for all assets with a LB = 0. 
* Turnover: maximum of 100% (which could be ignored but is shown for illustratation)
* RelWtsEquity: Sum of equity weights should be within 10% of benchmarket weights
* RelWtsUSFixed: Sum of US Fixed Income and Cash should be within 10% of benchmarket weights
* RelWtsOther: Sum of other assets should be within 10% of benchmarket weights
* WtsUSFixed: Sum of US Fixed Income and Cash should be at least 10%
* USLRelUS: US large should be at least 75% of US Large + US Small equity.  This type of constraint requires a bit of creativity. We want:
 - USL / (USL + USS) >= 0.75  (where USL is US Large and USS is US Small)
 - USL >= 0.75USL + 0.75USS so
 - 0.25USL - 0.75USS >= 0.  
 for this constraint we create a vector of coefficients (weights) with +0.25 for USL, -0.75 for USS, and 0 for the others.

##  The following are not part of the objective definition but will be needed:    
* x: Initial weights. Must sum to whatever value is desired (e.g., 1.0)
* LB: Lower bound weights for each asset
* UB: Upper bound weights for each asset

# Defining portfolio functions of x
These functions are included in the package.  User defined / custom functions may be used as well. The first parameter should
be x, a vector of portfolio weights.    
* calc_port_return 
* calc_port_risk
* calc_port_TE
* calc_nnonzeroassets
* calc_weighted
* calc_turnover
* calc_min_nonzero_weight

The following loads and creates some data which will be used in the example
```{r createData}
assign("data", rafi_201811)

# create a benchmark
benchX <- rep(0, nrow(data$expected.returns))
names(benchX) <- data$expected.returns$`Asset Class`
benchX[c('US Large', 'EAFE', 'Emerging Markets', 'US Aggregate')] <- c(0.36, 0.15, 0.09, 0.4)
benchX

# Lower and upper bounds
LB <- rep(0, nrow(data$expected.returns))
names(LB) <- data$expected.returns$`Asset Class`
LB["United States Cash"] <- .02

UB <- rep(1, nrow(data$expected.returns))
names(LB) <- data$expected.returns$`Asset Class`

# Current weights for calculating turnover
curX <- rep(0, nrow(data$expected.returns))
names(curX) <- data$expected.returns$`Asset Class`
curX["United States Cash"] <- 1

# definition of some groups
names.equity <- c("US Large", "US Small", "All country", "EAFE", "Emerging Markets", "Commodities", "REITS",
                 "US Commercial Real Estate")
names.fixed <- c("US Aggregate", "US Treasury Intermediate", "US Treasury Long", "US Treasury Short",
                "US Corporate Intermediate", "US Tips", "United States Cash") 
names.other <- c("Global Aggregate", "Global Ex-US Treasury", "Emerging Market (Local)", "Emerging Market (Non-Local)",
                "Global Ex-US Corporates", "Bank Loans", "US High Yield", "EM Cash", "Global DM Ex-US Long/Short Equity",
                "US Long/Short Equity", "Europe LBO", "US LBO") 
wts.equity <- numeric(nrow(data$expected.returns))
wts.equity[which(data$expected.returns$`Asset Class` %in% names.equity)] <- 1
wts.fixed <- numeric(nrow(data$expected.returns))
wts.fixed[which(data$expected.returns$`Asset Class` %in% names.fixed)] <- 1
wts.other <- numeric(nrow(data$expected.returns))
wts.other[which(data$expected.returns$`Asset Class` %in% names.other)] <- 1

wts.USLRelUS <- rep(0, nrow(data$expected.returns))
names(wts.USLRelUS) <- data$expected.returns$`Asset Class`
wts.USLRelUS[c('US Large', 'US Small')] <- c(0.25, -0.7)
```

# Build objective functions    
The following function `addTermToObjective` creates a list which defines a function, but does not create a function.  
The `calc_objective` function takes x and an objective definition and returns the value of the objective. Thus `calc_objective` can be the objective function if you don't want to supply one.

## Example: Using `addTermToObjective` to create a function definition    
```{r}
obj_definition <- addTermToObjective(name = "MaxReturn", 
                               type="maximize", 
                               funName=calc_port_return,
                               params = list(data$expected.returns$`Expected Return (Nominal)`)) %>%
            addTermToObjective(name = "Risk",
                               type = "constraint",
                               funName = calc_port_risk,
                               params = list(as.matrix(data$expected.cov[,-1])),
                               min = -Inf,
                               max = calc_port_risk(benchX, as.matrix(data$expected.cov[,-1]))) %>%
            addTermToObjective(name = "TrackingError",
                               type = "constraint",
                               funName = calc_port_TE,
                               params = list(benchX, as.matrix(data$expected.cov[,-1])),
                               min = -Inf,
                               max = 0.03) %>%
            addTermToObjective(name = "NAssets",
                               type = "constraint",
                               funName = calc_nnonzeroassets,
                               params = list(),
                               min = 3,
                               max = 10) %>%
           addTermToObjective(name = "MinNonZeroWt",
                              type = "constraint",
                              funName = calc_min_nonzero_weight,
                              params = list(LB),
                              min = 0.03,
                              max = Inf) %>%
          addTermToObjective(name = "Turnover",
                              type = "constraint",
                              funName = calc_turnover,
                              params = list(curX),
                              min = -Inf,
                              max = 1.0) %>%
          addTermToObjective(name = "RelWtsEquity",
                              type = "constraint",
                              funName = calc_weighted,
                              params = list(wts.equity),
                              min = sum(benchX[wts.equity == 1]) - 0.1,
                              max = sum(benchX[wts.equity == 1]) + 0.1) %>%
          addTermToObjective(name = "RelWtsFixed",
                              type = "constraint",
                              funName = calc_weighted,
                              params = list(wts.fixed),
                              min = sum(benchX[wts.fixed == 1]) - 0.1,
                              max = sum(benchX[wts.fixed == 1]) + 0.1) %>%
          addTermToObjective(name = "RelWtsOther",
                              type = "constraint",
                              funName = calc_weighted,
                              params = list(wts.other),
                              min = sum(benchX[wts.other == 1]) - 0.1,
                              max = sum(benchX[wts.other == 1]) + 0.1) %>%
          addTermToObjective(name = "WtsFixed",
                              type = "constraint",
                              funName = calc_weighted,
                              params = list(wts.fixed),
                              min = 0.1,
                              max = Inf)  %>%
          addTermToObjective(name = "USLRelUS",
                              type = "constraint",
                              funName = calc_weighted,
                              params = list(wts.USLRelUS),
                              min = 0,
                              max = Inf) 
  
```

## Examples: Using `calc_objective` to calculate the objective    
First we calculate assuming we invest benchmark weights    
```{r}
calc_objective(benchX, obj_definition = obj_definition)
```

The `calc_objective_terms` creates a tibble with each terms' value:
```{r}
calc_objective_terms(benchX, obj_definition)
```
Let's do the same thing with a portfolio where the penalties will be non-zero.  This would be curX.    
```{r}
calc_objective(curX, obj_definition = obj_definition)
```

```{r}
calc_objective_terms(curX, obj_definition)
```

# Optimizing    
Now let's optimize

```{r}
result <- maximize_objective(benchX, LB, UB, calc_objective, obj_definition=obj_definition)
result
```

Let's look at the results of the optimizer in the same way.    

```{r}
calc_objective(result$x, obj_definition = obj_definition)
```

```{r}
calc_objective_terms(result$x, obj_definition)
```
