---
title: "portOptimizer Example: Build objective function from scratch"
author: "Rex Macey"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro to portOptimizer building the objective function from scratch.}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(knitr)
library(kableExtra)
library(portOptimizer)
```



`portOptimizer` was created to handle complex portfolio optimization with problems of moderate size (number of assets <=100).  

Efficient portfolios provide the highest return for a given level of risk and satisfy other constraints.  `portOptimizer` is designed to provide flexibility regarding the complexity of the problem.  This example illustrates the flexibility of `portOptimizer`: It maximizes return subject to the following constraints:

- Minimum and Maximum weights in each asset class.    
- Maximum risk (absolute) is specified by an expected risk.  This might be either a number or the expected risk of a benchmark. 
- Maximum tracking error (relative risk) relative to a benchmark.
- Minimum and Maximum number assets.  
- Maximum turnover is the difference from the weightings of the optimal portfolio and the current holdings.  Selling all current holdings to buy new holdings would be a turnover of 100%.
- Minimum non-zero weight controls the minimum for a position size if there's a non-zero weight to the asset class. For example, perhaps if you are going to invest in an asset you want to allocate a minimum of 3% to it; otherwise it's a nuisance.  This constraint is ignored for asset classes with a minimum below this setting.  So if you indicate you want at least 2% in Cash, it will allow a 2% Cash position even if the Minimum non-zero weight parameter is 3%. 
- User defined constraints allow great flexibility to control the relative weights of one or more assets in the portfolio.  For example, a US investor with a domestic bias might require US Stocks to be at least 40% of all stocks in the portfolio.  These are specified when setting the cma.
- Class group constraints allow the weighting of one or more assets to be within a range of a benchmark weight.  For example, the portfolio equities might be constrainted to be within +/- 10% of the benchmark so that if the benchmark is 60% the portfolio's equity may range from 50% to 70%.

# The objective function

The main work on the user's part is to create an R function that takes in a set of asset weights (x) and returns a value.  portOptimizer tries to find the x that maximizes the value returned by the function.  In general, the objective function will look something like this:

calculate_my_objective_value <- function(x, expectedReturns, covmat, benchx, maxsigma, maxtracking, minnassets, maxnassets){
  port_return  <- sum(x * expectedReturns)
  port_sigma <- (max(0, t(x) %*% cma$cov %*% x)) ^ 0.5
  bench_sigma <- (max(0, t((x - benchx)) %*% cma$cov %*% (x - benchx))) ^ 0.5
  nnonzeroassets <- x != 0
  
  penalties <-  (ifelse(port_sigma <= maxsigma, 0, (port_sigma - maxsigma)^2) +
                ifelse(bench_sigma <= maxtracking, 0, (bench_sigma - maxtracking)^2) + 
                ifelse(nnonzeroassets > minnassets, 0, (nnonzeroassets - minnassets)^2) +
                ifelse(nnonzeroassets > maxnassets, 0, (nnonzeroassets - maxnassets)^2)) *
                25000000
  out <- port_return - penalties
  return(out)
}

Important things to note about the objective function:
- the first paramater, x is a set of weights for the portfolio.  It should sum to 1 or whatever value you want the sum of the weights to be.
- it returns a number.
- it has as many other parameters as you like.
- the constraints of the problem are treated as penalties.  If a constraint is satisfied, it contributes zero to the penalty. If non-zero it should be squared to produce a positive value.  Penalties are multiplied by a very large value (we use 25 billion). 

# Data

For this example, we use a set of assumptions developed by a third-party, [Research Affiliates](https://interactive.researchaffiliates.com/asset-allocation#!/?currency=USD&model=ER&scale=LINEAR&terms=NOMINAL) as of November 2018. This dataset is included with this package.  It contains 10 year forecasts for 27 asset classes.  Below we show a bit.  

```{r loaddata}
assign("data", rafi_201811)
```

```{r showdata, echo=FALSE}
kable(data$expected.returns[1:10,c(1,3,4,5,6,7)], digits = 3, caption = "Expected Returns")
kable(data$expected.cov[1:10,1:10], digits = 3, caption = "Expected Covariance")
```

# Creating an objective function from scratch

Rather than write one really big objective function, we will break it into smaller functions.  We will create one for the expected return and each constraint.  Then we will combine them into our objective function.  In calculating the objective function below, notice how each penalty has a square term and the entire value is multiplied by penaltyFactor which will be large.  There is no need for objective function to include a penalty if the weights violate lower or upper bounds (called box constraints) for each class. Lower and upper bounds on asset classes are supplied to the optimization function. 

The objective function must have a parameter x which are the portfolio weights. 

```{r createobjectivefunction}
calculate_my_objective_value <- function(x, expectedReturns, expectedCov, benchX, optParams){
  port_return <- calc_port_return(x, expectedReturns)
  port_risk <- calc_port_risk(x, expectedCov) 
  port_TE <- calc_port_risk(x - benchX, expectedCov)
  nnonzeroassets <- calc_nnonzeroassets(x)
  
  penalties <- (ifelse(port_risk <= optParams$maxRisk, 0, (port_risk - optParams$maxRisk)^2)  + 
               ifelse(port_TE <= optParams$maxTE, 0, (port_TE - optParams$maxTE)^2)  + 
               ifelse(nnonzeroassets >= optParams$minNAssets, 0, 
                      (nnonzeroassets - optParams$minNAssets)^2) +
               ifelse(nnonzeroassets <= optParams$maxNAssets, 0, 
                      (nnonzeroassets - optParams$maxNAssets)^2)) * optParams$penaltyFactor
  return(port_return - penalties)
}
```

Below we show the results of these functions.   For this illustration, we'll choose an equally weighted portfolio (x).  For tracking error, we will use a benchmark weights (benchX) with US Large = 0.36, EAFE = 0.15, Emerging Markets = 0.09 and US Aggregate = 0.40.  We will limit the portfolio risk to the risk of the benchmark, the tracking error to 0.03, and require at least 3 assets but no more than 10.

```{r createDataForFunctions}
x <- rep(1/nrow(data$expected.returns), nrow(data$expected.returns))
LB <- rep(0, nrow(data$expected.returns)) # lower bounds for weights
UB <- rep(1, nrow(data$expected.returns)) # upper bounds
benchX <- rep(0, nrow(data$expected.returns))
names(benchX) <- data$expected.returns$`Asset Class`
benchX[c('US Large', 'EAFE', 'Emerging Markets', 'US Aggregate')] <- c(0.36, 0.15, 0.09, 0.4)
optParams <- list(maxTE = 0.03, minNAssets = 3, maxNAssets = 10, LB=LB, UB=UB, 
                  penaltyFactor = 25e9)
# We use the risk of the benchmark as the maxRisk paramater
optParams$maxRisk <- calc_port_risk(benchX, as.matrix(data$expected.cov[,-1]))

calc_port_return(x, data$expected.returns$`Expected Return (Nominal)`)
calc_port_risk(x, as.matrix(data$expected.cov[,-1])) # cov needs to be a matrix, not a d.f.
calc_port_risk(x - benchX, as.matrix(data$expected.cov[,-1])) # This is tracking error
calc_nnonzeroassets(x)

# We use the risk of the benchmark as the maxRisk paramater
optParams$maxRisk <- calc_port_risk(benchX, as.matrix(data$expected.cov[,-1]))
```

Except for the number of assets, all these values are within our constraints.  Because the number of assets is higher than our maximum, we see a large negative value for the objective function. 

```{r objectiveExample1}
# This is the value of the objective function
calculate_my_objective_value(x, data$expected.returns$`Expected Return (Nominal)`, 
                             as.matrix(data$expected.cov[,-1]), 
                             benchX, optParams)
```

Let's change the portfolio weights to the benchmark weights and recalculate.  The risk of the portfolio will be the risk of the benchmark; the tracking error should be zero, and the number of nonzero assets will be 4 which is within our constraints. Thus, the total penalty should be zero and the objective value should equal the expected return. 

```{r}
x <- benchX
calc_port_return(x, data$expected.returns$`Expected Return (Nominal)`)
calc_port_risk(x, as.matrix(data$expected.cov[,-1])) # cov needs to be a matrix, not a d.f.
calc_port_risk(benchX - x, as.matrix(data$expected.cov[,-1])) # This is tracking error
calc_nnonzeroassets(x)
calculate_my_objective_value(x, data$expected.returns$`Expected Return (Nominal)`, 
                             as.matrix(data$expected.cov[,-1]), 
                             benchX, optParams)
```

That's the hard part - creating an objective to maximize and which produces very low values when constraints are not met.  

```{r runOptimizer}
result <- maximize_objective(x, optParams$LB, optParams$UB, calculate_my_objective_value,
                           expectedReturns = data$expected.returns$`Expected Return (Nominal)`,
                           expectedCov = as.matrix(data$expected.cov[,-1]),
                           benchX = benchX, optParams = optParams)
```

The result of the `maximize_objective function` is a list with two elements:    
* x are the best portfolio weights
* MaxObj is the value of the objective functions at x.  In our example, if MaxObj equals the expected return of the portfolio then we can infer there were no penalties and all constraints were met.

```{r showResult}
paste("Objective value: ", result$MaxObj)
cat("Vector of weights:")
result$x

paste("Return of Optimal Portfolio: ", calc_port_return(result$x, data$expected.returns$`Expected Return (Nominal)`))
```


